---
title: "iOS App Usage Data Exploration by Study Day"
author: "Gaming Reduction Study Analysis"
format: 
  html:
    page-layout: full
    code-fold: true
    code-summary: "Show code"
    toc: true
    toc-depth: 3
    toc-location: left
    grid:
      sidebar-width: 250px
      body-width: 1400px
      margin-width: 100px
    css: |
      .quarto-container {
        max-width: none !important;
      }
      .content {
        max-width: 1400px !important;
        margin: 0 auto;
      }
      table {
        font-size: 0.85em;
      }
      .cell-output-display {
        overflow-x: auto;
      }
editor: visual
---

## Overview

This notebook explores iOS app usage data from the gaming reduction study, focusing on data availability across study days (1-28) for each participant. The analysis uses enrollment dates to calculate study days and assess whether donated data exists for each day, with particular emphasis on gaming app usage patterns.

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(knitr)
library(DT)

# Set options
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Data Refresh Pipeline

```{r data-refresh}
# Set to FALSE to skip data refresh and use existing files
REFRESH_DATA <- FALSE

# Cache duration: refresh data if older than 1 hour
CACHE_DURATION_HOURS <- 1

# Define the output files to check for caching
FINAL_OUTPUT_FILE <- "../../downloads/diary_images/ios/aggregated_participant_data_enriched.csv"
CONTACT_LIST_FILE <- "../../.tmp/contact_list_with_embedded.csv"

# Check if we need to refresh data based on file age
should_refresh_data <- function(file_path, cache_hours = 1) {
  if (!file.exists(file_path)) {
    return(TRUE)
  }
  
  file_age_hours <- as.numeric(difftime(Sys.time(), file.mtime(file_path), units = "hours"))
  return(file_age_hours > cache_hours)
}

# Check if either main data or contact list needs refresh
needs_main_refresh <- should_refresh_data(FINAL_OUTPUT_FILE, CACHE_DURATION_HOURS)
needs_contact_refresh <- should_refresh_data(CONTACT_LIST_FILE, CACHE_DURATION_HOURS)

if (REFRESH_DATA && (needs_main_refresh || needs_contact_refresh)) {
  cat("=== REFRESHING iOS DATA FROM QUALTRICS ===\n")
  cat("Running complete iOS data pipeline...\n\n")
  
  # Step 1: Download Qualtrics images
  cat("Step 1/6: Downloading Qualtrics survey responses and images...\n")
  step1_result <- tryCatch({
    system2(
      "python3",
      args = c("../../monitoring/qualtrics_image_downloader.py"),
      stdout = TRUE,
      stderr = TRUE,
      wait = TRUE
    )
  }, error = function(e) {
    cat("❌ Error in Step 1:", e$message, "\n")
    return(NULL)
  })
  
  if (!is.null(attr(step1_result, "status")) && attr(step1_result, "status") != 0) {
    cat("⚠️ Warning: Step 1 exited with non-zero status\n")
    if (length(step1_result) > 0) {
      cat("Output:\n", paste(tail(step1_result, 10), collapse = "\n"), "\n")
    }
  } else {
    cat("✅ Step 1 completed successfully\n")
  }
  
  # Step 2: OCR analysis with Gemini
  cat("\nStep 2/6: Analyzing screenshots with Gemini OCR...\n")
  step2_result <- tryCatch({
    system2(
      "python3",
      args = c("../../monitoring/ocr/gemini_screenshot_analyzer.py"),
      stdout = TRUE,
      stderr = TRUE,
      wait = TRUE
    )
  }, error = function(e) {
    cat("❌ Error in Step 2:", e$message, "\n")
    return(NULL)
  })
  
  if (!is.null(attr(step2_result, "status")) && attr(step2_result, "status") != 0) {
    cat("⚠️ Warning: Step 2 exited with non-zero status\n")
    if (length(step2_result) > 0) {
      cat("Output:\n", paste(tail(step2_result, 10), collapse = "\n"), "\n")
    }
  } else {
    cat("✅ Step 2 completed successfully\n")
  }
  
  # Step 3: Aggregate participant data
  cat("\nStep 3/6: Aggregating participant data...\n")
  step3_result <- tryCatch({
    system2(
      "python3",
      args = c("../../monitoring/ocr/participant_aggregator.py"),
      stdout = TRUE,
      stderr = TRUE,
      wait = TRUE
    )
  }, error = function(e) {
    cat("❌ Error in Step 3:", e$message, "\n")
    return(NULL)
  })
  
  if (!is.null(attr(step3_result, "status")) && attr(step3_result, "status") != 0) {
    cat("⚠️ Warning: Step 3 exited with non-zero status\n")
    if (length(step3_result) > 0) {
      cat("Output:\n", paste(tail(step3_result, 10), collapse = "\n"), "\n")
    }
  } else {
    cat("✅ Step 3 completed successfully\n")
  }
  
  # Step 4: Convert to CSV
  cat("\nStep 4/6: Converting summary to CSV...\n")
  step4_result <- tryCatch({
    system2(
      "python3",
      args = c("../../monitoring/ocr/summary_to_csv.py"),
      stdout = TRUE,
      stderr = TRUE,
      wait = TRUE
    )
  }, error = function(e) {
    cat("❌ Error in Step 4:", e$message, "\n")
    return(NULL)
  })
  
  if (!is.null(attr(step4_result, "status")) && attr(step4_result, "status") != 0) {
    cat("⚠️ Warning: Step 4 exited with non-zero status\n")
    if (length(step4_result) > 0) {
      cat("Output:\n", paste(tail(step4_result, 10), collapse = "\n"), "\n")
    }
  } else {
    cat("✅ Step 4 completed successfully\n")
  }
  
  # Step 5: Classify gaming apps
  cat("\nStep 5/6: Classifying gaming applications...\n")
  step5_result <- tryCatch({
    system2(
      "python3",
      args = c("../../monitoring/ocr/app_game_classifier.py"),
      stdout = TRUE,
      stderr = TRUE,
      wait = TRUE
    )
  }, error = function(e) {
    cat("❌ Error in Step 5:", e$message, "\n")
    return(NULL)
  })
  
  if (!is.null(attr(step5_result, "status")) && attr(step5_result, "status") != 0) {
    cat("⚠️ Warning: Step 5 exited with non-zero status\n")
    if (length(step5_result) > 0) {
      cat("Output:\n", paste(tail(step5_result, 10), collapse = "\n"), "\n")
    }
  } else {
    cat("✅ Step 5 completed successfully\n")
  }
  
  # Step 6: Pull contact list data from Qualtrics
  cat("\nStep 6/6: Pulling contact list data from Qualtrics...\n")
  step6_result <- tryCatch({
    system2(
      "python3",
      args = c("../../monitoring/pull_contact_list.py", "--output", "contact_list_with_embedded"),
      stdout = TRUE,
      stderr = TRUE,
      wait = TRUE
    )
  }, error = function(e) {
    cat("❌ Error in Step 6:", e$message, "\n")
    return(NULL)
  })
  
  if (!is.null(attr(step6_result, "status")) && attr(step6_result, "status") != 0) {
    cat("⚠️ Warning: Step 6 exited with non-zero status\n")
    if (length(step6_result) > 0) {
      cat("Output:\n", paste(tail(step6_result, 10), collapse = "\n"), "\n")
    }
  } else {
    cat("✅ Step 6 completed successfully\n")
  }
  
  cat("\n🎉 iOS data pipeline completed!\n\n")
  
} else if (REFRESH_DATA) {
  file_age_hours <- as.numeric(difftime(Sys.time(), file.mtime(FINAL_OUTPUT_FILE), units = "hours"))
  cat("=== USING CACHED iOS DATA ===\n")
  cat(sprintf("Data file is %.1f hours old (cache valid for %d hours)\n", file_age_hours, CACHE_DURATION_HOURS))
  cat("Skipping refresh to save time. Set REFRESH_DATA <- FALSE to always skip refresh.\n\n")
} else {
  cat("=== USING EXISTING iOS DATA FILES ===\n")
  cat("Skipping data refresh (REFRESH_DATA = FALSE)\n\n")
}

# Check if output files exist and show their info
files_to_check <- c(FINAL_OUTPUT_FILE, CONTACT_LIST_FILE)
for (file in files_to_check) {
  if (file.exists(file)) {
    file_info <- file.info(file)
    cat(sprintf("✓ %s exists (modified: %s, size: %s bytes)\n", 
                basename(file), 
                file_info$mtime, 
                format(file_info$size, big.mark = ",")))
  } else {
    cat(sprintf("✗ %s not found - data pipeline may have failed\n", basename(file)))
  }
}
cat("\n")
```

## Data Loading

```{r load-data}
# Load the iOS aggregated app usage data
ios_data <- read_csv("../../downloads/diary_images/ios/aggregated_participant_data_enriched.csv")

# Load contact list data with embedded fields (Condition, Platforms, phoneType, EnrollmentDate)
if (file.exists("../../.tmp/contact_list_with_embedded.csv")) {
  contact_list_data <- read_csv("../../.tmp/contact_list_with_embedded.csv") %>%
    select(RANDOM_ID, Condition, Platforms, phoneType, EnrollmentDate) %>%
    filter(!is.na(RANDOM_ID) & RANDOM_ID != "") %>%
    distinct(RANDOM_ID, .keep_all = TRUE) %>%
    mutate(
      RANDOM_ID = as.character(RANDOM_ID),
      EnrollmentDate = ymd(EnrollmentDate)
    )
  
  # Check if any iOS participants are missing from contact list
  participants_without_condition <- ios_data %>%
    distinct(PID) %>%
    filter(!PID %in% contact_list_data$RANDOM_ID)
  
  if (nrow(participants_without_condition) > 0) {
    cat("⚠️ WARNING: Found", nrow(participants_without_condition), "participants without contact list data:\n")
    print(participants_without_condition$PID)
    cat("\nConsider re-running with REFRESH_DATA <- TRUE to pull fresh contact list data.\n\n")
  }
  
  enrollment_data <- contact_list_data
  
} else {
  cat("⚠️ WARNING: Contact list file not found. Using fallback from daily survey file.\n")
  cat("Consider re-running with REFRESH_DATA <- TRUE to pull contact list data.\n\n")
  
  # Fallback: Load enrollment data from the daily survey file
  enrollment_data <- read_csv("../../downloads/diary_images/HFF Gaming Reduction 3 - Daily Survey.csv") %>%
    # Skip the first 3 header rows and select relevant columns
    slice(-(1:3)) %>%
    select(RANDOM_ID, EnrollmentDate, Condition) %>%
    filter(!is.na(RANDOM_ID) & RANDOM_ID != "" & !is.na(EnrollmentDate)) %>%
    distinct(RANDOM_ID, .keep_all = TRUE) %>%
    mutate(
      RANDOM_ID = as.character(RANDOM_ID),
      EnrollmentDate = ymd(EnrollmentDate),
      # Add placeholder columns for consistency
      Platforms = NA_character_,
      phoneType = NA_character_
    )
}

# Display data structure
cat("iOS App Usage Data Structure:\n")
glimpse(ios_data)

cat("\nEnrollment/Contact List Data Structure:\n")
glimpse(enrollment_data)

cat("\nNumber of participants in iOS data:", n_distinct(ios_data$PID), "\n")
cat("Number of participants with enrollment/contact data:", nrow(enrollment_data), "\n")

# Show condition distribution if available
if ("Condition" %in% names(enrollment_data)) {
  condition_counts <- enrollment_data %>%
    count(Condition) %>%
    mutate(Condition = ifelse(is.na(Condition), "Missing", Condition))
  
  cat("\nCondition distribution in contact list:\n")
  print(condition_counts)
}
```

## Data Preparation

```{r data-prep}
# Process iOS data and join with enrollment information
ios_processed <- ios_data %>%
  mutate(
    PID = as.character(PID),
    Date = ymd(Date),
    Duration = as.numeric(Duration),
    DeviceType = tolower(trimws(DeviceType)),  # Standardize device type
    ProbGame = case_when(
      ProbGame == "Yes" ~ TRUE,
      ProbGame == "No" ~ FALSE,
      TRUE ~ NA
    )
  ) %>%
  # Join with enrollment data (now includes Condition, Platforms, phoneType)
  left_join(enrollment_data, by = c("PID" = "RANDOM_ID")) %>%
  filter(!is.na(EnrollmentDate)) %>%
  mutate(
    # Calculate study day based on enrollment date
    study_day = as.numeric(Date - EnrollmentDate) + 1
  ) %>%
  filter(study_day >= 1 & study_day <= 28)  # Focus on 28-day study period

# Check if any participants have missing Condition after join
participants_missing_condition <- ios_processed %>%
  filter(is.na(Condition)) %>%
  distinct(PID)

if (nrow(participants_missing_condition) > 0) {
  cat("\n⚠️ WARNING: After joining,", nrow(participants_missing_condition), "participants still missing Condition data:\n")
  print(participants_missing_condition$PID)
  cat("These participants may not be in the contact list yet.\n")
}

# Check for participants without enrollment data
missing_enrollment <- ios_data %>%
  filter(!PID %in% enrollment_data$RANDOM_ID) %>%
  distinct(PID)

if(nrow(missing_enrollment) > 0) {
  cat("Participants in iOS data without enrollment information:\n")
  print(missing_enrollment$PID)
} else {
  cat("All iOS participants have enrollment data.\n")
}

# Device type summary
device_summary <- ios_processed %>%
  group_by(DeviceType) %>%
  summarise(
    participants = n_distinct(PID),
    sessions = n(),
    total_duration = sum(Duration, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(participants))

cat("\nDevice Type Summary:\n")
print(device_summary)

cat("\nData after processing:\n")
cat("Participants with enrollment dates:", n_distinct(ios_processed$PID), "\n")
cat("Date range:", range(ios_processed$Date, na.rm = TRUE), "\n")
cat("Study day range:", range(ios_processed$study_day, na.rm = TRUE), "\n")
cat("Available device types:", paste(unique(ios_processed$DeviceType), collapse = ", "), "\n")
```

## Participant Overview

```{r participant-overview}
# Summary by participant with device type and contact list information
participant_summary <- ios_processed %>%
  group_by(PID, Condition, Platforms, phoneType, EnrollmentDate) %>%
  summarise(
    device_types = paste(sort(unique(DeviceType)), collapse = ", "),
    total_apps = n_distinct(App),
    total_sessions = n(),
    total_duration = sum(Duration, na.rm = TRUE),
    gaming_apps = n_distinct(App[ProbGame == TRUE]),
    gaming_sessions = sum(ProbGame == TRUE, na.rm = TRUE),
    gaming_duration = sum(Duration[ProbGame == TRUE], na.rm = TRUE),
    study_days_with_data = n_distinct(study_day),
    first_day = min(study_day),
    last_day = max(study_day),
    .groups = "drop"
  ) %>%
  mutate(
    # Calculate study completion status
    report_generation_date = Sys.Date(),
    study_end_date = EnrollmentDate + days(27), # 28 days total (0-27)
    days_since_enrollment = as.numeric(report_generation_date - EnrollmentDate),
    study_should_be_complete = days_since_enrollment >= 28,
    study_completion_status = case_when(
      !study_should_be_complete ~ "🟡 In Progress",
      study_should_be_complete & study_days_with_data >= 20 ~ "✅ Complete",
      days_since_enrollment > 30 & study_days_with_data < 20 ~ "🔴 Incomplete",
      TRUE ~ "❓ Unknown"
    ),
    status_category = case_when(
      !study_should_be_complete ~ "in_progress",
      study_should_be_complete & study_days_with_data >= 20 ~ "complete",
      days_since_enrollment > 30 & study_days_with_data < 20 ~ "incomplete",
      TRUE ~ "unknown"
    ),
    gaming_percentage = round((gaming_duration / total_duration) * 100, 1)
  ) %>%
  arrange(status_category, PID)

# Display interactive participant table
datatable(participant_summary %>% select(-status_category), 
      caption = "iOS Participant Summary with Study Completion Status, Contact List Info, and Device Types",
      colnames = c("Participant ID", "Condition", "Study Platforms", "Phone Type", "Enrollment Date", "Device Types", "Total Apps", "Total Sessions", 
                   "Total Duration (min)", "Gaming Apps", "Gaming Sessions", "Gaming Duration (min)",
                   "Days with Data", "First Day", "Last Day", "Report Date", "Study End Date", 
                   "Days Since Enrollment", "Should Complete", "Status", "Gaming %"),
      options = list(
        pageLength = 20,
        scrollX = TRUE,
        columnDefs = list(list(width = '120px', targets = c(0, 1, 2, 3, 19))),
        order = list(list(19, 'asc'))
      ),
      filter = 'top') %>%
      formatRound(columns = c(7, 10, 20), digits = 1)

# Summary of completion status
status_summary <- participant_summary %>%
  count(study_completion_status, status_category) %>%
  arrange(status_category)

cat("\n=== iOS STUDY COMPLETION STATUS SUMMARY ===\n")
for(i in 1:nrow(status_summary)) {
  cat(status_summary$study_completion_status[i], ":", status_summary$n[i], "participants\n")
}
```

## Data Availability Matrix

```{r data-availability-matrix}
# Create comprehensive data availability matrix for iOS participants by device type
study_day_availability <- ios_processed %>%
  group_by(PID, DeviceType, study_day) %>%
  summarise(
    has_data = TRUE,
    num_apps = n_distinct(App),
    total_duration = sum(Duration, na.rm = TRUE),
    gaming_duration = sum(ifelse(ProbGame == TRUE & !is.na(ProbGame), Duration, 0), na.rm = TRUE),
    .groups = "drop"
  )

# Debug: Check gaming duration calculation by device type
cat("Gaming duration check by device type:\n")
gaming_by_device <- study_day_availability %>%
  group_by(DeviceType) %>%
  summarise(
    total_gaming_duration = sum(gaming_duration, na.rm = TRUE),
    participants = n_distinct(PID),
    .groups = "drop"
  )
print(gaming_by_device)

# Let's also check the raw data by device type
gaming_check_device <- ios_processed %>%
  filter(ProbGame == TRUE) %>%
  group_by(DeviceType) %>%
  summarise(
    total_gaming_sessions = n(),
    total_gaming_duration = sum(Duration, na.rm = TRUE),
    .groups = "drop"
  )
cat("Raw gaming data check by device type:\n")
print(gaming_check_device)

# Get all participants and their device types
participant_devices <- ios_processed %>%
  group_by(PID) %>%
  summarise(device_types = paste(sort(unique(DeviceType)), collapse = ", "), .groups = "drop")

# Create separate complete grids for each device type
device_types <- unique(ios_processed$DeviceType)
all_study_days <- 1:28

# Create device-specific grids
device_grids <- list()

for(device in device_types) {
  device_participants <- unique(ios_processed$PID[ios_processed$DeviceType == device])
  
  device_grid <- expand_grid(
    PID = device_participants,
    study_day = all_study_days,
    DeviceType = device
  ) %>%
    left_join(study_day_availability, by = c("PID", "DeviceType", "study_day")) %>%
    mutate(
      has_data = ifelse(is.na(has_data), FALSE, has_data),
      gaming_duration = ifelse(is.na(gaming_duration), 0, gaming_duration)
    ) %>%
    left_join(
      participant_summary %>% select(PID, Condition, EnrollmentDate, 
                                   study_completion_status, status_category),
      by = "PID"
    )
  
  device_grids[[device]] <- device_grid
  
  cat(sprintf("Device '%s': %d participants, %d total records\n", 
              device, length(device_participants), nrow(device_grid)))
}

# Combine all device grids
complete_grid <- bind_rows(device_grids)

# Debug: Check the join results
cat("\nCombined grid dimensions:", nrow(complete_grid), "rows\n")
cat("Participants with Condition data:", sum(!is.na(complete_grid$Condition)), "\n")
cat("Unique participants in complete_grid:", n_distinct(complete_grid$PID), "\n")

# Summary statistics by device type
availability_summary_by_device <- complete_grid %>%
  group_by(DeviceType, study_day) %>%
  summarise(
    participants_with_data = sum(has_data),
    total_participants = n(),
    percentage_with_data = (participants_with_data / total_participants) * 100,
    .groups = "drop"
  )

# Overall summary (combined)
availability_summary <- complete_grid %>%
  group_by(study_day) %>%
  summarise(
    participants_with_data = sum(has_data),
    total_participants = n(),
    percentage_with_data = (participants_with_data / total_participants) * 100,
    .groups = "drop"
  )

# Display device-specific summaries
for(device in device_types) {
  device_summary <- availability_summary_by_device %>% filter(DeviceType == device)
  
  cat(sprintf("\n=== %s DATA AVAILABILITY SUMMARY ===\n", toupper(device)))
  device_table <- datatable(device_summary %>% select(-DeviceType),
            caption = paste("Data Availability Summary by Study Day -", stringr::str_to_title(device)),
            colnames = c("Study Day", "Participants with Data", "Total Participants", "Percentage with Data"),
            options = list(
              pageLength = 10,
              scrollX = TRUE,
              order = list(list(0, 'asc'))
            ),
            filter = 'top') %>%
    formatRound(columns = c(4), digits = 1)
  
  print(device_table)
}
```

## Visualization: Data Availability Heatmaps by Device Type

```{r heatmap-visualization}
# Create separate heatmaps for each device type
for(device in device_types) {
  device_data <- complete_grid %>% filter(DeviceType == device)
  device_participants <- unique(device_data$PID)
  
  cat(sprintf("\n=== %s DEVICE HEATMAP ===\n", toupper(device)))
  cat(sprintf("Participants: %d\n", length(device_participants)))
  
  if(nrow(device_data) > 0) {
    # Create heatmap for this device type
    device_heatmap <- device_data %>%
      # Add completion status colors to participant labels
      mutate(
        participant_label = paste0(PID, " (", str_extract(study_completion_status, "^[🟡✅🔴❓]"), ")")
      ) %>%
      # Sort by completion status for better visualization
      arrange(status_category, PID) %>%
      mutate(participant_label = factor(participant_label, levels = unique(participant_label))) %>%
      ggplot(aes(x = study_day, y = participant_label, fill = has_data)) +
      geom_tile(color = "white", size = 0.1) +
      scale_fill_manual(
        values = c("FALSE" = "#f7f7f7", "TRUE" = "#2166ac"),
        labels = c("No Data", "Has Data"),
        name = "Data Status"
      ) +
      scale_x_continuous(breaks = seq(1, 28, 2)) +
      labs(
        title = paste("App Usage Data Availability by Study Day -", stringr::str_to_title(device)),
        x = "Study Day",
        y = "Participant ID (Completion Status)",
        caption = paste0("Blue = Data available, Gray = No data. ", stringr::str_to_title(device), " device only (", 
                        length(device_participants), " participants).\nParticipants sorted by completion status: 🟡 In Progress, ✅ Complete, 🔴 Incomplete.")
      ) +
      theme_minimal() +
      theme(
        axis.text.y = element_text(size = 8),
        axis.text.x = element_text(size = 10),
        plot.title = element_text(size = 14, face = "bold"),
        plot.subtitle = element_text(size = 11),
        legend.position = "bottom"
      )
    
    print(device_heatmap)
  } else {
    cat("No data available for this device type.\n")
  }
}

# Combined heatmap with device type faceting
combined_heatmap <- complete_grid %>%
  # Add completion status colors to participant labels
  mutate(
    participant_label = paste0(PID, " (", str_extract(study_completion_status, "^[🟡✅🔴❓]"), ")"),
    DeviceType = stringr::str_to_title(DeviceType)
  ) %>%
  # Sort by completion status for better visualization
  arrange(DeviceType, status_category, PID) %>%
  mutate(participant_label = factor(participant_label, levels = unique(participant_label))) %>%
  ggplot(aes(x = study_day, y = participant_label, fill = has_data)) +
  geom_tile(color = "white", size = 0.1) +
  scale_fill_manual(
    values = c("FALSE" = "#f7f7f7", "TRUE" = "#2166ac"),
    labels = c("No Data", "Has Data"),
    name = "Data Status"
  ) +
  scale_x_continuous(breaks = seq(1, 28, 4)) +
  facet_wrap(~ DeviceType, scales = "free_y", ncol = 1) +
  labs(
    title = "App Usage Data Availability by Study Day and Device Type",
    x = "Study Day",
    y = "Participant ID (Completion Status)",
    caption = "Blue = Data available, Gray = No data. Participants sorted by completion status within each device type."
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 7),
    axis.text.x = element_text(size = 10),
    plot.title = element_text(size = 14, face = "bold"),
    strip.text = element_text(size = 12, face = "bold"),
    legend.position = "bottom"
  )

cat("\n=== COMBINED DEVICE TYPE HEATMAP ===\n")
print(combined_heatmap)
```

## Gaming Time Series Analysis

```{r gaming-timeseries}
# Calculate daily gaming duration for each participant by device type
gaming_timeseries <- complete_grid %>%
  mutate(
    gaming_duration = ifelse(is.na(gaming_duration), 0, gaming_duration),
    gaming_duration_min = gaming_duration  # Data is already in minutes
  )

# Only filter out participants without Condition data if they exist, otherwise keep all
if(sum(is.na(gaming_timeseries$Condition)) < nrow(gaming_timeseries)) {
  gaming_timeseries <- gaming_timeseries %>%
    filter(!is.na(Condition))
  cat("Filtered to participants with Condition data\n")
} else {
  cat("Warning: All participants missing Condition data - keeping all participants\n")
}

# Check gaming data by device type
gaming_by_device_summary <- gaming_timeseries %>%
  group_by(DeviceType) %>%
  summarise(
    total_gaming_time = sum(gaming_duration_min, na.rm = TRUE),
    gaming_entries = sum(gaming_duration_min > 0, na.rm = TRUE),
    participants = n_distinct(PID),
    .groups = "drop"
  )

cat("Gaming time summary by device type:\n")
print(gaming_by_device_summary)

# Overall gaming check
total_gaming_time <- sum(gaming_timeseries$gaming_duration_min, na.rm = TRUE)
cat("\nTotal gaming time across all participants and devices:", total_gaming_time, "minutes\n")

# Show sample of gaming data by device type
for(device in device_types) {
  device_gaming <- gaming_timeseries %>%
    filter(DeviceType == device & gaming_duration_min > 0) %>%
    head(5)
  
  if(nrow(device_gaming) > 0) {
    cat(sprintf("\nSample %s gaming entries:\n", device))
    print(device_gaming %>% select(PID, DeviceType, study_day, gaming_duration_min, Condition))
  }
}

# Only create plots if we have gaming data
if(total_gaming_time > 0) {
  # Create separate time series plots for each device type
  for(device in device_types) {
    device_gaming_data <- gaming_timeseries %>% filter(DeviceType == device)
    device_gaming_time <- sum(device_gaming_data$gaming_duration_min, na.rm = TRUE)
    
    cat(sprintf("\n=== %s GAMING TIME SERIES ===\n", toupper(device)))
    
    if(device_gaming_time > 0) {
      # Find participants with gaming data for this device
      participants_with_gaming_device <- device_gaming_data %>%
        group_by(PID) %>%
        summarise(total_gaming = sum(gaming_duration_min, na.rm = TRUE)) %>%
        filter(total_gaming > 0) %>%
        pull(PID)
      
      cat(sprintf("Participants with %s gaming data: %d\n", device, length(participants_with_gaming_device)))
      
      if(length(participants_with_gaming_device) > 0) {
        # Prepare data for plotting
        gaming_plot_data <- device_gaming_data %>%
          filter(PID %in% participants_with_gaming_device) %>%
          mutate(
            plot_condition = ifelse(is.na(Condition), "Unknown", as.character(Condition)),
            plot_color = case_when(
              is.na(Condition) ~ PID,
              TRUE ~ as.character(Condition)
            )
          )
        
        # Check for NA conditions
        has_na_conditions <- any(is.na(gaming_plot_data$Condition))
        
        if(has_na_conditions) {
          na_participants <- unique(gaming_plot_data$PID[is.na(gaming_plot_data$Condition)])
          base_colors <- c("control" = "#d73027", "intervention" = "#1a9850")
          
          if(length(na_participants) > 0) {
            na_colors <- rainbow(length(na_participants), start = 0.3, end = 0.8)
            names(na_colors) <- na_participants
            all_colors <- c(base_colors, na_colors)
          } else {
            all_colors <- base_colors
          }
          
          device_gaming_plot <- gaming_plot_data %>%
            ggplot(aes(x = study_day, y = gaming_duration_min, color = plot_color)) +
            geom_line(aes(group = PID), alpha = 0.7, size = 0.8) +
            {if(any(!is.na(gaming_plot_data$Condition))) {
              geom_smooth(data = gaming_plot_data %>% filter(!is.na(Condition)), 
                         aes(color = Condition), method = "loess", se = TRUE, size = 1.5, alpha = 0.7)
            }} +
            scale_x_continuous(breaks = seq(1, 28, 2)) +
            scale_color_manual(values = all_colors, name = "Condition/Participant") +
            labs(
              title = paste("Gaming Duration Time Series -", stringr::str_to_title(device)),
              x = "Study Day",
              y = "Gaming Duration (minutes)",
              caption = paste0("Individual trajectories for ", stringr::str_to_title(device), " device. Based on ", length(participants_with_gaming_device), " participants.")
            ) +
            theme_minimal() +
            theme(legend.position = "bottom")
          
        } else {
          device_gaming_plot <- gaming_plot_data %>%
            ggplot(aes(x = study_day, y = gaming_duration_min, color = Condition)) +
            geom_line(aes(group = PID), alpha = 0.3, size = 0.5) +
            geom_smooth(method = "loess", se = TRUE, size = 1.5, alpha = 0.7) +
            scale_x_continuous(breaks = seq(1, 28, 2)) +
            scale_color_manual(values = c("control" = "#d73027", "intervention" = "#1a9850")) +
            labs(
              title = paste("Gaming Duration Time Series -", stringr::str_to_title(device)),
              x = "Study Day",
              y = "Gaming Duration (minutes)",
              color = "Condition",
              caption = paste0("Individual trajectories (thin lines) with smoothed averages (thick lines) for ", stringr::str_to_title(device), " device.\nBased on ", length(participants_with_gaming_device), " participants.")
            ) +
            theme_minimal() +
            theme(legend.position = "bottom")
        }
        
        print(device_gaming_plot)
      } else {
        cat(sprintf("No participants found with %s gaming data.\n", device))
      }
    } else {
      cat(sprintf("No gaming time recorded for %s devices.\n", device))
    }
  }
  
  # Combined plot with device type faceting
  participants_with_gaming <- gaming_timeseries %>%
    group_by(PID, DeviceType) %>%
    summarise(total_gaming = sum(gaming_duration_min, na.rm = TRUE), .groups = "drop") %>%
    filter(total_gaming > 0)
  
  if(nrow(participants_with_gaming) > 0) {
    cat("\n=== COMBINED GAMING TIME SERIES BY DEVICE TYPE ===\n")
    
    combined_gaming_plot <- gaming_timeseries %>%
      filter(paste(PID, DeviceType) %in% paste(participants_with_gaming$PID, participants_with_gaming$DeviceType)) %>%
      mutate(DeviceType = stringr::str_to_title(DeviceType)) %>%
      ggplot(aes(x = study_day, y = gaming_duration_min, color = Condition)) +
      geom_line(aes(group = PID), alpha = 0.4, size = 0.5) +
      geom_smooth(method = "loess", se = TRUE, size = 1.2, alpha = 0.7) +
      scale_x_continuous(breaks = seq(1, 28, 4)) +
      scale_color_manual(values = c("control" = "#d73027", "intervention" = "#1a9850")) +
      facet_wrap(~ DeviceType, scales = "free_y", ncol = 1) +
      labs(
        title = "Gaming Duration Time Series by Device Type and Condition",
        x = "Study Day",
        y = "Gaming Duration (minutes)",
        color = "Condition",
        caption = "Individual participant trajectories with smoothed condition averages, separated by device type."
      ) +
      theme_minimal() +
      theme(
        legend.position = "bottom",
        strip.text = element_text(size = 12, face = "bold")
      )
    
    print(combined_gaming_plot)
  }
} else {
  cat("No gaming data found. Skipping gaming time series plots.\n")
}
```

## Gaming Apps Analysis

```{r gaming-apps-analysis}
# Identify and analyze gaming apps
gaming_apps_summary <- ios_processed %>%
  filter(ProbGame == TRUE) %>%
  group_by(App) %>%
  summarise(
    total_sessions = n(),
    total_duration = sum(Duration, na.rm = TRUE),
    unique_users = n_distinct(PID),
    unique_days = n_distinct(Date),
    avg_duration_per_session = mean(Duration, na.rm = TRUE),
    avg_confidence = mean(LLM_conf, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(unique_users >= 2) %>%  # Apps used by at least 2 participants
  arrange(desc(total_duration)) %>%
  mutate(
    total_duration_hours = round(total_duration / 60, 1),
    avg_duration_per_session = round(avg_duration_per_session, 1),
    avg_confidence = round(avg_confidence, 1)
  )

datatable(gaming_apps_summary,
          caption = "Gaming Apps Summary (Used by 2+ Participants)",
          colnames = c("App Name", "Total Sessions", "Total Duration (min)", "Unique Users", 
                       "Days Used", "Avg Duration/Session (min)", "Avg LLM Confidence", 
                       "Total Duration (hrs)"),
          options = list(
            pageLength = 15,
            scrollX = TRUE,
            columnDefs = list(list(width = '150px', targets = c(0))),
            order = list(list(2, 'desc'))
          ),
          filter = 'top') %>%
          formatRound(columns = c(2, 5, 6, 7), digits = 1)

# Top gaming apps by participant
top_gaming_by_participant <- ios_processed %>%
  filter(ProbGame == TRUE) %>%
  group_by(PID, App) %>%
  summarise(
    total_sessions = n(),
    total_duration = sum(Duration, na.rm = TRUE),
    days_used = n_distinct(Date),
    .groups = "drop"
  ) %>%
  group_by(PID) %>%
  arrange(PID, desc(total_duration)) %>%
  mutate(
    app_rank = row_number(),
    pct_of_total_gaming = round((total_duration / sum(total_duration)) * 100, 1)
  ) %>%
  filter(app_rank <= 5) %>%  # Top 5 gaming apps per participant
  ungroup() %>%
  left_join(
    participant_summary %>% select(PID, Condition, study_completion_status),
    by = "PID"
  )

# Show sample of top gaming apps by participant
cat("Sample of participants' top gaming apps:\n")
sample_participants <- top_gaming_by_participant %>%
  arrange(PID, app_rank)

current_participant <- ""
count <- 0
for(i in 1:min(nrow(sample_participants), 50)) {  # Show first 50 entries
  row <- sample_participants[i,]
  if (row$PID != current_participant) {
    current_participant <- row$PID
    count <- count + 1
    if(count <= 10) {  # Show first 10 participants
      cat(sprintf("\n%s (%s, %s):\n", row$PID, row$study_completion_status, row$Condition))
    }
  }
  if(count <= 10) {
    cat(sprintf("  %d. %s (%d sessions, %.1f min, %d days, %.1f%%)\n", 
                row$app_rank, row$App, row$total_sessions, row$total_duration, 
                row$days_used, row$pct_of_total_gaming))
  }
}
```

## Data Availability Trends

```{r availability-trends}
# Plot percentage of participants with data by study day
availability_trend <- availability_summary %>%
  ggplot(aes(x = study_day, y = percentage_with_data)) +
  geom_line(color = "#2166ac", size = 1) +
  geom_point(color = "#2166ac", size = 2) +
  scale_x_continuous(breaks = seq(1, 28, 2)) +
  scale_y_continuous(limits = c(0, 100), labels = function(x) paste0(x, "%")) +
  labs(
    title = "Percentage of iOS Participants with App Usage Data by Study Day",
    x = "Study Day",
    y = "Percentage of Participants with Data",
    caption = paste0("Based on ", n_distinct(ios_processed$PID), " iOS participants")
  ) +
  theme_minimal()

print(availability_trend)

# Additional summary statistics
cat("\nData Availability Statistics:\n")
cat("- Days with 100% participation:", sum(availability_summary$percentage_with_data == 100), "\n")
cat("- Days with >50% participation:", sum(availability_summary$percentage_with_data > 50), "\n")
cat("- Days with <25% participation:", sum(availability_summary$percentage_with_data < 25), "\n")
```

## Individual Participant Patterns

```{r individual-patterns}
# Look at individual participant data patterns
participant_patterns <- complete_grid %>%
  group_by(PID, Condition, study_completion_status, status_category) %>%
  summarise(
    total_days_with_data = sum(has_data),
    consecutive_days = max(rle(has_data)$lengths[rle(has_data)$values == TRUE], na.rm = TRUE),
    first_day_with_data = min(study_day[has_data], na.rm = TRUE),
    last_day_with_data = max(study_day[has_data], na.rm = TRUE),
    total_gaming_duration = sum(gaming_duration, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    consecutive_days = ifelse(is.infinite(consecutive_days), 0, consecutive_days),
    data_span = last_day_with_data - first_day_with_data + 1,
    avg_gaming_per_day = round(total_gaming_duration / total_days_with_data, 1)
  ) %>%
  # Sort by completion status
  arrange(status_category, PID)

# Summary by condition
pattern_summary <- participant_patterns %>%
  group_by(Condition) %>%
  summarise(
    n_participants = n(),
    avg_days_with_data = mean(total_days_with_data, na.rm = TRUE),
    median_days_with_data = median(total_days_with_data, na.rm = TRUE),
    avg_consecutive_days = mean(consecutive_days, na.rm = TRUE),
    avg_gaming_duration = mean(total_gaming_duration, na.rm = TRUE),
    .groups = "drop"
  )

cat("iOS Participant Data Patterns by Condition:\n")
datatable(pattern_summary,
          caption = "iOS Participant Data Patterns Summary by Condition",
          colnames = c("Condition", "Participants", "Avg Days with Data", "Median Days with Data", 
                       "Avg Consecutive Days", "Avg Total Gaming Duration (min)"),
          options = list(
            pageLength = 10,
            scrollX = TRUE,
            order = list(list(0, 'asc'))
          ),
          filter = 'top') %>%
          formatRound(columns = c(3, 4, 5, 6), digits = 1)

# Interactive detailed participant table
datatable(participant_patterns, 
      caption = "Individual iOS Participant Data Patterns (Sorted by Completion Status)",
      options = list(
        pageLength = 20,
        scrollX = TRUE,
        columnDefs = list(list(width = '100px', targets = c(0, 1, 2, 3))),
        order = list(list(2, 'asc'))
      ),
      filter = 'top') %>%
      formatRound(columns = c(5, 6, 7, 8, 9, 10, 11), digits = 1)
```

## Key Findings Summary

```{r summary}
# Calculate key statistics
total_ios_participants <- n_distinct(ios_processed$PID)
days_with_any_data <- sum(availability_summary$participants_with_data > 0)
peak_participation_day <- availability_summary$study_day[which.max(availability_summary$percentage_with_data)]
peak_participation_rate <- max(availability_summary$percentage_with_data)

avg_days_per_participant <- mean(participant_patterns$total_days_with_data, na.rm = TRUE)
participants_with_full_study <- sum(participant_patterns$total_days_with_data >= 20, na.rm = TRUE)

# Gaming statistics
total_gaming_apps <- n_distinct(ios_processed$App[ios_processed$ProbGame == TRUE])
avg_gaming_duration <- mean(participant_patterns$total_gaming_duration, na.rm = TRUE)

# Completion status counts
completion_counts <- participant_summary %>% count(study_completion_status)
in_progress <- completion_counts$n[completion_counts$study_completion_status == "🟡 In Progress"]
complete <- completion_counts$n[completion_counts$study_completion_status == "✅ Complete"]
incomplete <- completion_counts$n[completion_counts$study_completion_status == "🔴 Incomplete"]

# Handle NA values for missing statuses
in_progress <- ifelse(length(in_progress) == 0, 0, in_progress)
complete <- ifelse(length(complete) == 0, 0, complete)
incomplete <- ifelse(length(incomplete) == 0, 0, incomplete)

cat("=== iOS APP USAGE KEY FINDINGS ===\n\n")
cat("Study Overview:\n")
cat("- Total iOS participants with app usage data:", total_ios_participants, "\n")
cat("- Report generated on:", as.character(Sys.Date()), "\n")
cat("- Study days with any data:", days_with_any_data, "out of 28\n")
cat("- Peak participation on day", peak_participation_day, "with", round(peak_participation_rate, 1), "% of participants\n\n")

cat("Study Completion Status:\n")
cat("- 🟡 In Progress (not yet 28 days since enrollment):", in_progress, "participants\n")
cat("- ✅ Complete (≥28 days, ≥20 days of data):", complete, "participants\n")
cat("- 🔴 Incomplete (≥28 days, <20 days of data):", incomplete, "participants\n\n")

cat("Data Coverage:\n")
cat("- Average days with data per participant:", round(avg_days_per_participant, 1), "\n")
cat("- Participants with data for ≥20 days:", participants_with_full_study, "out of", total_ios_participants, "\n")
cat("- Overall data availability rate:", round(mean(availability_summary$percentage_with_data), 1), "%\n\n")

cat("Gaming Analysis:\n")
cat("- Unique gaming apps identified:", total_gaming_apps, "\n")
cat("- Average total gaming duration per participant:", round(avg_gaming_duration, 1), "minutes\n")
cat("- Gaming apps used by ≥2 participants:", nrow(gaming_apps_summary), "\n\n")

cat("Condition Comparison:\n")
control_avg <- pattern_summary$avg_days_with_data[pattern_summary$Condition == "control"]
intervention_avg <- pattern_summary$avg_days_with_data[pattern_summary$Condition == "intervention"]
control_gaming <- pattern_summary$avg_gaming_duration[pattern_summary$Condition == "control"]
intervention_gaming <- pattern_summary$avg_gaming_duration[pattern_summary$Condition == "intervention"]

if(length(control_avg) > 0) cat("- Control group average days:", round(control_avg, 1), "\n")
if(length(intervention_avg) > 0) cat("- Intervention group average days:", round(intervention_avg, 1), "\n")
if(length(control_gaming) > 0) cat("- Control group average gaming duration:", round(control_gaming, 1), "min\n")
if(length(intervention_gaming) > 0) cat("- Intervention group average gaming duration:", round(intervention_gaming, 1), "min\n")
```

## Data Export

```{r export}
# Create output directory if it doesn't exist
if (!dir.exists("../../.tmp/ios")) {
  dir.create("../../.tmp/ios", recursive = TRUE)
}

# Export processed data for further analysis
write_csv(complete_grid, "../../.tmp/ios/study_day_availability_matrix.csv")
write_csv(availability_summary, "../../.tmp/ios/daily_availability_summary.csv")
write_csv(participant_patterns, "../../.tmp/ios/participant_data_patterns.csv")
write_csv(gaming_timeseries, "../../.tmp/ios/gaming_timeseries_data.csv")
write_csv(gaming_apps_summary, "../../.tmp/ios/gaming_apps_summary.csv")
write_csv(top_gaming_by_participant, "../../.tmp/ios/top_gaming_apps_by_participant.csv")

cat("Exported iOS analysis files:\n")
cat("- study_day_availability_matrix.csv: Complete participant x study day matrix\n")
cat("- daily_availability_summary.csv: Daily availability percentages\n")
cat("- participant_data_patterns.csv: Individual participant patterns\n")
cat("- gaming_timeseries_data.csv: Gaming duration time series data\n")
cat("- gaming_apps_summary.csv: Gaming apps usage statistics\n")
cat("- top_gaming_apps_by_participant.csv: Top gaming apps per participant\n")
```