---
title: "H2 - Effect of Abstention on Sleep Quality"
output: html_document
execute-dir: project
format: 
  html:
    toc: true
    code-fold: true
author: "Tamás Földes"
date: today
---

## Introduction

This document demonstrates a complete workflow for generating synthetic data, introducing dropout, and fitting models (a GAM or MLM) to the data. The main focus is on the `sim_study` function, which orchestrates the data simulation, dropout process, and model fitting.

As an overview, we compare the following possible models for estimating the effect of gaming reduction, for both H2a (the intention-to-treat effect) and H2b (the per-protocol effect; i.e., the effect of actual gaming reduction relative to one's own baseline). The models we compare are:

| Model Name | Syntax | Target Effect | Notes |
|------------|----------------------|------------|---------------------------|
| GAM | `gam(sleep_quality ~ condition:intervention_period + age + gender + s(id, bs = "re") + s(day, by = condition, bs = "tp"), correlation = corAR1(form = ~ day | id))` | ITT |  |
| GAM with no main effect | `gam(sleep_quality ~         age + gender +         s(id, bs = "re") +         s(day, by = condition, bs = "tp"),       data = dat,       correlation = corCAR1(form = ~ day | id))` | ITT | Here we do not estimate a parameter for the effect of the intervention directly; rather, we simply fit separate curves to each condition and calculate the average marginal effect using {emmeans} |
| MLM | `lme(     fixed = sleep_quality ~ condition*intervention_period + age + gender,     random = ~ 1|id,     correlation = corCAR1(form = ~ day | id),     method = "ML"   )` | ITT | Multiple versions of this model failed when including random slopes; we therefore dropped these |
| MLM Simple | `lme(     fixed = sleep_quality ~ baseline + condition + age + gender,     random = ~ 1 + condition|id,     correlation = corCAR1(form = ~ day | id),     method = "ML",   )` | ITT | Here we do not model the baseline (pre-intervention period) itself—we model only the 14-day period when the intervention is active, using average sleep_quality during baseline as a covariate |
| GLS (generalized least squares) | `gls(     sleep_quality ~ condition * intervention_period + age + gender,     correlation = corCAR1(form = ~ day | id),   )` | ITT |  |
| GLS Simple | `gls(     sleep_quality ~ condition + baseline + age + gender,     correlation = corAR1(form = ~ day | id),   )` | ITT |  |
| GLS Splines | `gls(     sleep_quality ~ ns(day, df = 4) * intervention_period * condition,     correlation = corCAR1(form = ~ day | id),     data = dat   )` | ITT | In this version, we fit a GLS but allow non-linearity in the trajectory of sleep_quality using splines |
| MLM Reduction | `lme(     fixed = sleep_quality ~ intervention_active*reduction + age + gender,     random = ~ 1 + intervention_active*reduction | id,     correlation = corCAR1(form = ~ day | id)   )` | PP | Here we test our intended model for the per-protocol effect; `reduction` is the number of hours played relative to that person's mean playtime at baseline |

: ITT = Intention-to-treat; PP = per-protocol

### Take-aways

Our simulations show that several models perform well at parameter recovery for the ITT effect, but that the GAM model has the highest power for small effects---the type of effects we believe we are most likely to observed---and for non-linear trajectories over the 14 day period (e.g., an effect that slowly accumulates over a couple of days and then plateaus, or a temporary withdrawal followed by a later improvement). The GAM has approximately 50% power for a standardized effect of .2, and 80% power for a standardized effect of .3, but this varies based the shape of that effect over time. 

The MLM Reduction model performs very well, and has >95% power for standardized effects of approximately .2 or greater. 

### Load Libraries

First we load packages with `pacman`, which is fully compatible with `renv`.

```{r}
#| label: load-libraries
#| code-summary: "Show code (load libraries)"

library(pacman)

p_load(tidyverse, qualtRics, lme4, mgcv, marginaleffects, broom, forestplot, broom.mixed, nlme, rms, emmeans, splines, furrr, extraDistr)

# Replace your current plan() line with:
if (interactive()) {
  plan(multisession, workers = parallel::detectCores()-8)
} else {
  plan(sequential)
}
# Diagnostic information about parallel setup

message("=== PARALLEL PROCESSING SETUP ===")
message("Total CPU cores detected: ", parallel::detectCores())
message("Number of workers allocated: ", nbrOfWorkers())
message("Future plan: ", paste(class(plan()), collapse = ", "))
message("===================================")

theme_set(theme_minimal())
theme_update(
  strip.background = element_rect(fill = "black"),
  strip.text = element_text(color = "white", size = 10),
  panel.grid.minor = element_blank(),
  panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
)

options(scipen = 999)

```

### Simulation, Dropout, and Fitting Functions

Here we define:

-   `sim_data`: Generates synthetic data with random intercepts/slopes and AR(1) errors.
-   `sim_dropout`: Introduces missingness and dropout in the dataset.
-   `fit_*`: Fits a statistical model to the simulated data (see table above)
-   `sim_study`: Ties everything together—generates data, applies dropout, then fits the chosen model and returns a tidy summary.

```{r}
#| label: sim-functions
#| code-summary: "Show code (sim functions)"

#' Generate Synthetic Data
#'
#' This function simulates synthetic panel data for `n` participants over `n_days` time points,
#' with an intervention effect, random intercepts and slopes, and AR(1)-correlated residuals.
#'
#' @param n Number of participants. Default is 100.
#' @param n_days Number of time points per participant.
#' @param b Fixed effect of condition (if mediated == FALSE) or a 1-hour reduction in playtime.
#' @param phi AR(1) autocorrelation coefficient.
#' @param sigma AR(1) residual standard deviation.

sim_data <- function(n = 100,
                     n_days = 28,
                     
                     # effect parameters
                     b = 0.8, # effect in unstandardized units (scaled for 5-25 range)
                     mu = 15, # grand mean of the outcome (center of 5-25 range)
                     
                     
                     # random effects parameters
                     tau_int = 2.5,   # Random intercept SD (between-person variance)
                     tau_slope = .05,  # Random slope SD 
                     within_person_sd = 2.0, # Within-person SD (scaled for 5-25 range)
                     
                     # AR(1) parameters
                     phi = 0.8,     # Autocorrelation coefficient
                     effect_shape = "grow",
                     k = .5, # affects how quickly the plateau effect plateaus
                     
                     mediated = FALSE,
                     
                     # playtime parameters
                     playtime_grand_mean = 1,   # Average baseline playtime in hours
                     playtime_grand_sd = .5,   # SD for baseline playtime in log units (log-normal distribution)
                     daily_play_sd = 0.5      # Daily noise in playtime
                     # compliance_mean = 0.7,    # Average reduction (in hours) for intervention group during intervention period
)     
{
  dat <- tibble(
    id = 1:n,
    age = sample(18:36, n, replace = TRUE),
    gender = sample(c("man","woman","non-binary"), n, prob = c(.45, .45, .1), replace = TRUE),
    condition = factor(sample(c("control", "intervention"), n, replace = TRUE)),
    experimental_condition = ifelse(condition == "intervention", 1, 0),
    intercept_sq = rnorm(n, 0, tau_int), # Renamed from intercept_wb
    slope_sq = rnorm(n, 0, tau_slope), # Renamed from slope_wb
    intercept_play = rlnorm(n, log(playtime_grand_mean), playtime_grand_sd),
  ) |> 
    # expand to 28 waves per id
    crossing(
      day = 1:n_days
    ) |> 
    mutate(
      intervention_period = as.numeric(day > 7 & day < 22),
      intervention_active = intervention_period & condition == "intervention",
      compliance = ifelse(intervention_active, rkumar(n*n_days, a = .05, b = .1), 0),
      
      # In the baseline period, play is just the subject's baseline plus some day-to-day noise
      # During the intervention, experimental subjects reduce play by their compliance amount
      playtime = (1 - compliance) * rlnorm(n, log(intercept_play), daily_play_sd),
      effect_time = case_when(
        effect_shape == "plateau" ~ if_else(intervention_period == 1, (b + slope_sq) * (1-exp(-k * (day - 7))), 0), # Renamed from slope_wb
        effect_shape == "grow" ~ if_else(intervention_period == 1, (day - 7) * ((b + slope_sq)/7), 0), # Renamed from slope_wb
        TRUE ~ NA_real_
      ),
    ) |> 
    group_by(id) |> 
    mutate(
      
      baseline_playtime = mean(playtime[day <= 7]),
      reduction = baseline_playtime - playtime, # The mediator: reduction in play relative to the baseline average
      sigma = within_person_sd * sqrt(1-phi^2),
      # Generate AR(1) errors for each participant
      e = as.numeric(arima.sim(n = n_days, 
                               model = list(ar = phi), 
                               sd = sigma)),
      # Add random effect + fixed effect + AR(1) error
      sleep_quality_raw = case_when( # Renamed from wellbeing
        mediated == TRUE ~ mu + 
                            intercept_sq + # Renamed from intercept_wb
                            effect_time * reduction + 
                            .01*(age-18) +
                            -.05*(gender %in% c("women","non-binary")) + 
                            e,
        mediated == FALSE ~ mu + 
                            intercept_sq + # Renamed from intercept_wb
                            effect_time * experimental_condition * intervention_period + 
                            .01*(age-18) +
                            -.05*(gender %in% c("women","non-binary")) + 
                            e
      ),
      # Constrain to 5-25 range and round to integers
      sleep_quality = round(pmax(5, pmin(25, sleep_quality_raw)))
    ) |> 
    ungroup() |> 
    mutate(across(where(is.numeric), ~ round(., 3)))
  
  dat

}
```
```{r}
#| label: sim-data-example
#| code-summary: "Show code (sim data example)"

sim_data(n = 10, n_days = 10, effect_shape = "grow")


#' Simulate Dropout
#'
#' Introduces missingness and dropout into a dataset by randomly assigning records as missing
#' or dropped out. Once a participant is dropped out, all subsequent records become missing.
#'
#' @param dat A tibble generated by \code{sim_data()}.
#'
#' @return A tibble of the same structure as \code{dat}, but with some \code{sleep_quality} values set to NA. # Renamed from wellbeing
#'
sim_dropout <- function(dat) {
  
  dropout <- dat |> 
    mutate(
      missing = sample(c(TRUE, FALSE), n(), replace = TRUE, prob = c(.10, .90)),
      dropout = sample(c(TRUE, FALSE), n(), replace = TRUE, prob = c(.01, .99))
    ) |>
    mutate(
      missing = ifelse(cumsum(dropout) > 0, TRUE, missing),
      .by = id
    ) |>
    arrange(as.integer(id), day) |> 
    mutate(sleep_quality = ifelse(missing, NA, sleep_quality)) # Renamed from wellbeing
  dropout
}

```

```{r}
#| label: fit-functions
#| code-summary: "Show code (fit functions)"

#' Fit a Generalized Additive Model (GAM)
#'
#' Fits a GAM model to the provided dataset using \code{mgcv::gam}, including an AR(1)
#' correlation structure and random intercept for each ID.
#'
#' @param dat A tibble of repeated-measures data (e.g., from \code{sim_data()} and \code{sim_dropout()}).
#'
#' @return An object of class \code{gam}, which is the fitted GAM model.
#'
#' 
fit_gam <- function(dat) {
  
  gam(sleep_quality ~ # Renamed from wellbeing
        condition:intervention_period + age + gender +
        s(id, bs = "re") + 
        s(day, by = condition, bs = "tp"), 
      data = dat,
      correlation = corAR1(form = ~ day | id))
}

fit_gam_no_main <- function(dat) {
  
  gam(sleep_quality ~ # Renamed from wellbeing
        age + gender +
        s(id, bs = "re") + 
        s(day, by = condition, bs = "tp"), 
      data = dat,
      correlation = corCAR1(form = ~ day | id))
}

#' Fit a Multi-Level Model (MLM)
#'
#' Fits a linear mixed-effects model (LME) with random intercept for each ID using \code{lme4::lmer}.
#'
#' @param dat A tibble of repeated-measures data (e.g., from \code{sim_data()} and \code{sim_dropout()}).
#'
#' @return An object of class \code{lmerMod}, which is the fitted MLM model.
#'
fit_mlm <- function(dat) {
  # lmer(sleep_quality ~ condition*intervention_period + age + gender + (1|id), data = dat) # Renamed from wellbeing
  lme(
    fixed = sleep_quality ~ condition*intervention_period + age + gender, # Renamed from wellbeing
    random = ~ 1|id,  # or use a more flexible structure if needed
    correlation = corCAR1(form = ~ day | id),
    method = "ML",
    data = dat |> filter(!is.na(sleep_quality)) # Renamed from wellbeing
  )
}

fit_mlm_simple <- function(dat) {
  
  tmp <- dat |> 
    group_by(id) |> 
    # take the mean of days 1-7 
    mutate(baseline = mean(sleep_quality[day < 8], na.rm = TRUE)) |> # Renamed from wellbeing
    filter(intervention_period == 1) |> 
    filter(!is.na(sleep_quality)) # Renamed from wellbeing
  
  lme(
    fixed = sleep_quality ~ baseline + condition + age + gender, # Renamed from wellbeing
    random = ~ 1|id,  # Only random intercept since condition is between-subjects
    correlation = corCAR1(form = ~ day | id),
    method = "ML",
    data = tmp |> filter(!is.na(sleep_quality)) # Renamed from wellbeing
  )
}

fit_gls <- function(dat) {
  
  gls(
    sleep_quality ~ condition * intervention_period + age + gender, # Renamed from wellbeing
    correlation = corCAR1(form = ~ day | id),
    data = dat |> filter(!is.na(sleep_quality)) # Renamed from wellbeing
  )
}

fit_gls_simple <- function(dat) {
  
  tmp <- dat |> 
    group_by(id) |> 
    # take the mean of days 1-7 
    mutate(baseline = mean(sleep_quality[day < 8], na.rm = TRUE)) |> # Renamed from wellbeing
    filter(intervention_period == 1) |> 
    filter(!is.na(sleep_quality)) # Renamed from wellbeing
  
  gls(
    sleep_quality ~ condition + baseline + age + gender, # Renamed from wellbeing
    correlation = corAR1(form = ~ day | id),
    data = tmp
  )
}

fit_gls_spline <- function(dat) {
  gls(
    sleep_quality ~ ns(day, df = 4) * intervention_period * condition, # Renamed from wellbeing, corrected comma
    correlation = corCAR1(form = ~ day | id),
    data = dat
  )
}

fit_mlm_reduction <- function(dat) {
  lme(
    fixed = sleep_quality ~ intervention_active*reduction + age + gender, # Renamed from wellbeing
    random = ~ 1 | id,  # Simplified to random intercept only
    correlation = corCAR1(form = ~ day | id),
    data = dat
  )
}

# Helper function to extract the focal effect for GLS models
extract_marginal_effect <- function(mod, dat, focal_term = "conditionintervention") {
  # Here we assume your GLS model is specified with condition*intervention_period
  # and you want the effect of condition (e.g., intervention vs. control) during intervention.
  # We create a reference grid that fixes intervention_period at 1.
  rg <- ref_grid(mod, data = dat, at = list(intervention_period = 1))
  
  # Obtain estimated marginal means for each condition.
  emm <- emmeans(rg, ~ condition)
  
  # Compute the pairwise contrast (e.g., intervention - control)
  # Adjust names as needed. The contrast below returns a one-row summary.
  contr <- emmeans::contrast(emm, method = list("intervention - control" = c(-1, 1)), adjust = "none")
  contr_sum <- summary(contr, infer = TRUE)
  
  # Construct a one-row data frame with consistent column names.
  # If you have more than one contrast, you might need to filter for the one of interest.
  df <- data.frame(
    term = focal_term,
    estimate = contr_sum$estimate,
    std.error = contr_sum$SE,
    conf.low = contr_sum$lower.CL,
    conf.high = contr_sum$upper.CL,
    row.names = NULL
  )
  
  return(df)
}
```
```{r}
#' Simulation Study Orchestrator
#'
#' A higher-level function that ties together data simulation, dropout, and model fitting,
#' returning a tidy summary of the fitted model parameters.
#'
#' @param model_function A function to fit the model. Defaults to \code{fit_gam}.
#' @param n Number of participants passed to \code{sim_data()}. Default is 1000.
#' @param n_days Number of time points per participant passed to \code{sim_data()}. Default is 28.
#' @param b Fixed effect for the intervention slope passed to \code{sim_data()}. Default is 0.01.
#' @param phi AR(1) autocorrelation coefficient passed to \code{sim_data()}. Default is 0.7.
#' @param sigma AR(1) residual standard deviation passed to \code{sim_data()}. Default is 0.6.
#'
#' @return A data frame (tibble) of model estimates from \code{broom::tidy(parametric = TRUE)}.
#'

# Updated simulation orchestrator that handles GLS models separately.
sim_study <- function(model = "fit_gam", focal_term = "intervention_activeTRUE:reduction", ...) {
  args <- list(...)
  dat <- do.call(sim_data, args)
  model_function <- get(model)
  
  mod <- model_function(dat)
  
  if (model %in% c("fit_gam_no_main","fit_gls_spline")) {
    # Extract the effect using our helper function.
    result <- suppressMessages(extract_marginal_effect(mod, 
                                                       dat, 
                                                       focal_term = focal_term))
  } else {
    # For models that work with broom, extract the focal parameter.
    # Adjust the filtering term as needed.
    result <- broom::tidy(mod, parametric = TRUE) |>
      filter(term == focal_term) |> 
      # filter(
      #   term == "conditionintervention:intervention_period" | 
      #     (model %in% c("fit_mlm_simple","fit_gls_simple") & term == "conditionintervention")
      # ) |> 
      mutate(
        conf.low = estimate - 1.96 * std.error,
        conf.high = estimate + 1.96 * std.error
      )
  }
  result
}
```

## Test and Plot One Simulated Study

Below, we create a sample dataset using `sim_data()` and examine it with a line plots by day. We can also see whether the simulated SDs for wellbeing align with the target values in the simulation---luckily, they do.

```{r}
#| label: test-plot-descriptive
#| code-summary: "Show code (descriptive plotting)"

dat <- sim_data(effect_shape = "plateau", mediated = TRUE)

sds <- dat |> 
  group_by(id) |> 
  summarise(mean_value = mean(sleep_quality, na.rm = TRUE), # Renamed from wellbeing
            sd_within  = sd(sleep_quality, na.rm = TRUE)) |> # Renamed from wellbeing
  summarise(between_sd   = sd(mean_value, na.rm = TRUE),
            avg_within_sd = mean(sd_within, na.rm = TRUE))

# plot sleep_quality by group
dat |> 
  group_by(condition, day) |> 
  summarise(sleep_quality = mean(sleep_quality)) |> # Renamed from wellbeing
  ggplot(aes(y = sleep_quality, x = day, color = condition)) + # Renamed from wellbeing
  geom_line() + 
  theme_minimal()

```

### Test Fit (h2a - intention to treat)

We fit various models to the newly simulated data to make sure each appears to be working properly, and also test the full `sim_study` pipeline.

```{r}
#| label: test-fit-h2a
#| eval: false
#| code-summary: "Show code (test fit h2a)"

dat <- sim_data(mediated = FALSE)

fit_mlm(dat) |> summary()
fit_mlm_simple(dat) |> summary()
fit_gls(dat) |> summary()
fit_gls_simple(dat) |> summary()
fit_gls_spline(dat) |> extract_marginal_effect()
fit_gam(dat) |> summary()
fit_gam_no_main(dat) |> extract_marginal_effect()

sim_study(model = "fit_gam_no_main")
sim_study(model = "fit_gls_spline")

```

Since some models (e.g., `fit_gam_no_main`) do not have a parameter that represents the average difference-in-difference between groups during the intervention period, we need to calculate this ourselves by marginalize across the 14-day intervention period.

```{r}
#| label: test-emmeans
#| eval: false
#| code-summary: "Show code (test emmeans)"

emm_day <- emmeans(
  fit_gls_spline(dat), 
  pairwise ~ condition | day, 
  at = list(day = 8:21), 
  condition = c("control", "intervention"), 
  data = dat |> mutate(condition = factor(condition, levels = c("intervention", "control")))
)

summary(emm_day$contrasts, infer = TRUE, level = .95, by = NULL, adjust = "none")

# and then integrated over the 14 day intervention period
rg <- ref_grid(fit_gls_spline(dat),
               at = list(intervention_period = 1),
               cov.reduce = list(day = mean),
               data = dat |> mutate(condition = factor(condition, levels = c("control","intervention"))))

emm <- emmeans(rg, ~ condition)
(contrast_result <- contrast(emm, method = list("intervention - control" = c(-1, 1)), adjust = "none"))


means <- summary(emm)$emmean
names(means) <- summary(emm)$condition
(diff_manual <- means["intervention"] - means["control"])
```

### Test Fit (h2b - per-protocol)

Another quick test of our `fit_mlm_reduction` model, to make sure the alternative simulation whereby the effect of the intervention is mediated by a reduction in playtime is also functioning properly.

```{r}
#| label: test-fit-h2b
#| eval: false
#| code-summary: "Show code (test fit h2b)"

dat <- sim_data(mediated = TRUE)

fit_mlm_reduction(dat) |> summary()

```

## Simulated h2a power analysis

To assess power/sensitivity, we run multiple simulations (controlled by `n_sims`) and gather the parameter estimates for a particular term (e.g., `conditionintervention:intervention_periodTRUE`, or for our marginalized effect `conditionintervention`). Each iteration calls `sim_study`, which does the data generation, dropout, and fitting.

As this is quite slow, we both use parallel processing with `furrr` cache the results.

```{r, cache = TRUE}
#| label: sim-study-h2a
#| code-summary: "Show code (simulate power h2a)"

# Load tictoc for timing if not already loaded
if (!require(tictoc, quietly = TRUE)) {
  install.packages("tictoc")
  library(tictoc)
}

# Start timing
tic("H2a simulation total time")

n_sims <- 500

specs_h2a <- expand_grid(
  model = c("fit_gam", "fit_gam_no_main", "fit_mlm", "fit_mlm_simple", "fit_gls", "fit_gls_simple", "fit_gls_spline"),  # model names as strings
  b = c(0.3, 0.6, 0.9, 1.2, 1.5),  # scaled for 5-25 range
  effect_shape = c("grow", "plateau")
) |> 
  mutate(
    focal_term = case_when(
      model %in% c("fit_mlm_simple","fit_gls_simple", "fit_gam_no_main") ~ "conditionintervention",
      model %in% c("fit_gam", "fit_mlm", "fit_gls", "fit_gls_spline") ~ "conditionintervention:intervention_period",
      TRUE ~ "conditionintervention:intervention_period"
    )
  ) |> 
  (function(d) { d$row_id <- pmap_chr(d, ~ paste0(names(list(...)), "=", c(...), collapse = "_")); d })() |> 
  mutate(i = row_number())

# Create all spec-simulation combinations for better parallelization
all_jobs_h2a <- specs_h2a |> 
  crossing(sim = 1:n_sims) |> 
  mutate(job_id = row_number())

message("Running ", nrow(all_jobs_h2a), " simulations across ", nbrOfWorkers(), " cores...")

# Run all simulations in parallel
results_h2a <- future_map_dfr(1:nrow(all_jobs_h2a), function(job_idx) {
  # Load required libraries in each worker
  library(tidyverse)
  library(lme4)
  library(mgcv)
  library(broom)
  library(nlme)
  library(emmeans)
  library(splines)
  library(broom.mixed)
  library(extraDistr)
  
  # Get job parameters
  job <- all_jobs_h2a[job_idx, ]
  
  tryCatch({
    result <- sim_study(
      model = job$model,
      focal_term = job$focal_term,
      n = 100,
      n_days = 28,
      # effect parameters
      b = job$b,
      mu = 15,
      effect_shape = job$effect_shape,
      k = .5,
      # random effects parameters
      tau_int = 2.5,
      tau_slope = .8,
      within_person_sd = 2.5,
      # AR(1) parameters
      phi = 0.7
    ) |> 
      mutate(
        sim = job$sim,
        row_id = job$row_id,
        model = job$model,
        b = job$b,
        effect_shape = job$effect_shape
      )
    
    return(result)
  }, error = function(e) {
    message("Job ", job_idx, " (spec row: ", job$i, ", sim: ", job$sim, ") failed: ", e$message)
    tibble(
      term = NA_character_,
      estimate = NA_real_,
      std.error = NA_real_,
      conf.low = NA_real_,
      conf.high = NA_real_,
      sim = job$sim,
      row_id = job$row_id,
      model = job$model,
      b = job$b,
      effect_shape = job$effect_shape
    )
  })
}, .progress = TRUE,
.options = furrr_options(
  globals = c("all_jobs_h2a", "sim_study", "sim_data", 
              "fit_gam", "fit_gam_no_main",
              "fit_mlm", "fit_mlm_simple",
              "fit_gls", "fit_gls_simple",
              "fit_gls_spline", 
              "extract_marginal_effect"),
  seed = TRUE
))

sim_summary_h2a <- results_h2a |> 
  group_by(row_id) |> 
  summarise(
    model = first(model),
    b = first(b),
    effect_shape = first(effect_shape),
    mean_effect = mean(estimate, na.rm = TRUE),
    mean_se = mean(std.error, na.rm = TRUE),
    mean_conf.low = mean(conf.low, na.rm = TRUE),
    mean_conf.high = mean(conf.high, na.rm = TRUE),
    power = sum(conf.low > 0, na.rm = TRUE) / sum(!is.na(conf.low))
  )

# Stop timing and display results
toc()

# Also print some summary statistics about the simulation
message("=== SIMULATION SUMMARY ===")
message("Total number of jobs: ", nrow(all_jobs_h2a))
message("Number of successful results: ", sum(!is.na(results_h2a$estimate)))
message("Number of failed jobs: ", sum(is.na(results_h2a$estimate)))
message("Success rate: ", round(100 * sum(!is.na(results_h2a$estimate)) / nrow(results_h2a), 1), "%")
message("============================")


```

```{r}
#| label: visualize-power-h2a
#| code-summary: "Show code (visualize power h2a)"

# Estimated effect vs. true effect (b)
ggplot(sim_summary_h2a, aes(x = b, y = mean_effect, color = model, alpha = model == "fit_gam")) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_conf.low, ymax = mean_conf.high), width = 0.1) +
  facet_wrap(~ effect_shape) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  scale_alpha_manual(values = c("TRUE" = 1.0, "FALSE" = 0.2), guide = "none") +
  labs(x = "True Effect (unstandardized b)", y = "Estimated Effect",
       title = "Estimated vs. True Effects by Model and Effect Shape") +
  scale_x_continuous(breaks = c(0.3, 0.6, 0.9, 1.2, 1.5),
                     sec.axis = sec_axis(~ . / 3, name = "Standardized Effect (b/3)"))

# ggsave("figures/estimated_vs_true_effects.png", width = 10, height = 6, dpi = 300)

# Power vs. true effect (b)
ggplot(sim_summary_h2a, aes(x = b, y = power, color = model, alpha = model == "fit_gam")) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  facet_wrap(~ effect_shape) +
  scale_alpha_manual(values = c("TRUE" = 1.0, "FALSE" = 0.2), guide = "none") +
  labs(x = "True Effect (unstandardized b)", y = "Power",
       title = "Power by Model and Effect Shape") +
  scale_x_continuous(breaks = c(0.3, 0.6, 0.9, 1.2, 1.5),
                     sec.axis = sec_axis(~ . / 3, name = "Standardized Effect (b/3)")) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1), breaks = seq(0, 1, .1))

# ggsave("figures/power_by_model_and_effect_shape.png", width = 10, height = 6, dpi = 300)

# results |>
#   forestplot(mean = estimate,
#              lower = conf.low,
#              upper = conf.high,
#              labeltext = term)

```

## Simulated h2b power analysis

Same thing as above, but now looking at power for our per-protocol model.

```{r, cache = TRUE}
#| label: sim-study-h2b
#| code-summary: "Show code (sim study h2b)"


# Start timing
tic("H2b simulation total time")

n_sims <- 500

specs_h2b <- expand_grid(
  model = c("fit_mlm_reduction"),  # model names as strings
  b = c(0.5, 1.0, 1.5, 2.0, 2.5),  # these seem reasonable for the mediated model
  effect_shape = c("grow", "plateau")
) |> 
  # calculate the mean of the Kumaraswamy distribution - the expected effect size of the mediated version is b * average compliance
  mutate(
    focal_term = case_when(
      model == "fit_mlm_reduction" ~ "intervention_activeTRUE:reduction"
    ),
    expected_effect = b * .1*beta(1 + 1/.05, .1),
  ) |> 
  (\(d) { d$row_id <- pmap_chr(d, ~ paste0(names(list(...)), "=", c(...), collapse = "_")); d })() |> 
  mutate(i = row_number())

# Create all spec-simulation combinations for better parallelization
all_jobs_h2b <- specs_h2b |> 
  crossing(sim = 1:n_sims) |> 
  mutate(job_id = row_number())

message("Running ", nrow(all_jobs_h2b), " simulations across ", nbrOfWorkers(), " cores...")

# Run all simulations in parallel
results_h2b <- future_map_dfr(1:nrow(all_jobs_h2b), function(job_idx) {
  # Load required libraries in each worker
  library(tidyverse)
  library(nlme)
  library(broom.mixed)
  library(extraDistr)
  library(rms)
  
  # Get job parameters
  job <- all_jobs_h2b[job_idx, ]
  
  tryCatch({
    result <- sim_study(
      model = job$model,
      focal_term = job$focal_term,
      n = 100,
      n_days = 28,
      # effect parameters
      b = job$b,
      mu = 15,
      effect_shape = job$effect_shape,
      k = .5,
      # random effects parameters
      tau_int = 2.5,
      tau_slope = .8,
      within_person_sd = 2.0,
      # AR(1) parameters
      phi = 0.7,
      mediated = TRUE
    ) |> 
      mutate(
        sim = job$sim,
        row_id = job$row_id,
        model = job$model,
        b = job$b,
        expected_effect = job$expected_effect,
        effect_shape = job$effect_shape
      )
    
    return(result)
  }, error = function(e) {
    message("Job ", job_idx, " (spec row: ", job$i, ", sim: ", job$sim, ") failed: ", e$message)
    tibble(
      term = NA_character_,
      estimate = NA_real_,
      std.error = NA_real_,
      conf.low = NA_real_,
      conf.high = NA_real_,
      sim = job$sim,
      row_id = job$row_id,
      model = job$model,
      b = job$b,
      expected_effect = job$expected_effect,
      effect_shape = job$effect_shape
    )
  })
}, .progress = TRUE,
.options = furrr_options(
  globals = c("all_jobs_h2b", "sim_study", "sim_data", "fit_mlm_reduction"),
  seed = TRUE
))

sim_summary_h2b <- results_h2b |> 
  group_by(row_id) |> 
  summarise(
    model = first(model),
    b = first(b),
    expected_effect = first(expected_effect),
    effect_shape = first(effect_shape),
    mean_effect = mean(estimate, na.rm = TRUE),
    mean_se = mean(std.error, na.rm = TRUE),
    mean_conf.low = mean(conf.low, na.rm = TRUE),
    mean_conf.high = mean(conf.high, na.rm = TRUE),
    power = sum(conf.low > 0, na.rm = TRUE) / sum(!is.na(conf.low))
  )

# Stop timing and display results
toc()

# Also print some summary statistics about the simulation
message("=== H2B SIMULATION SUMMARY ===")
message("Total number of jobs: ", nrow(all_jobs_h2b))
message("Number of successful results: ", sum(!is.na(results_h2b$estimate)))
message("Number of failed jobs: ", sum(is.na(results_h2b$estimate)))
message("Success rate: ", round(100 * sum(!is.na(results_h2b$estimate)) / nrow(results_h2b), 1), "%")
message("===============================")


```

```{r}
#| label: visualize-power-h2b
#| code-summary: "Show code (visualize power h2b)"

# Estimated effect vs. true effect (b)
ggplot(sim_summary_h2b, aes(x = expected_effect, y = mean_effect, color = model)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_conf.low, ymax = mean_conf.high), width = 0.1) +
  facet_wrap(~ effect_shape) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  labs(x = "True Effect (expected effect)", y = "Estimated Effect",
       title = "Estimated vs. True Effects by Model and Effect Shape") +
  scale_x_continuous(breaks = c(0.05, 0.1, 0.15, 0.2, 0.25),
                     sec.axis = sec_axis(~ . / 3, name = "Standardized Effect"))

# ggsave("figures/estimated_vs_true_effects_h2b.png", width = 10, height = 6, dpi = 300)

# Power vs. true effect (b)
ggplot(sim_summary_h2b, aes(x = expected_effect, y = power, color = model)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  facet_wrap(~ effect_shape) +
  labs(x = "True Effect (expected effect)", y = "Power",
       title = "Power by Model and Effect Shape") +
  scale_x_continuous(breaks = c(0.05, 0.1, 0.15, 0.2, 0.25),
                     sec.axis = sec_axis(~ . / 3, name = "Standardized Effect")) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1), breaks = seq(0, 1, .1))

# ggsave("figures/power_by_model_and_effect_shape_h2b.png", width = 10, height = 6, dpi = 300)  
```

## Planned Sensitivity Analyses

We have preregistered several sensitivity analyses to test the robustness of any effects we find. These include:

- *Day 21 only:* We will estimate the effect of the intervention on day 21 only, to see what the difference between groups is at the end of the intervention period
- *Marginal means:* In h2a, we will use the `emmeans` package to calculate marginal means for each condition and produce a single parameter by integrating across the 14-day period
- *Multilevel model:* We will fit the MLM as defined in Table 1 above, as the second-highest performing model in the simulations (having higher power for linear effects, but lower for non-linear)