---
title: "Generate Synthetic Data"
output: html_document
execute-dir: project
---

## Introduction

This document demonstrates a complete workflow for generating synthetic data, introducing dropout, and fitting models (a GAM or MLM) to the data. The main focus is on the `sim_study` function, which orchestrates the data simulation, dropout process, and model fitting. The code is presented in separate chunks, each serving a specific purpose, with explanatory text interspersed.

First we load packages with `pacman`, which is fully compatible with `renv`.
```{r}
#| label: load-libraries

library(pacman)

p_load(tidyverse, qualtRics, lme4, mgcv, marginaleffects, broom, forestplot, broom.mixed, nlme)

theme_set(theme_minimal())
theme_update(
  strip.background = element_rect(fill = "black"),
  strip.text = element_text(color = "white", size = 10),
  panel.grid.minor = element_blank(),
  panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
)

```

### Simulation, Dropout, and Fitting Functions

Here we define:

- `sim_data`: Generates synthetic data with random intercepts/slopes and AR(1) errors.
- `sim_dropout`: Introduces missingness and dropout in the dataset.
- `fit_gam`: Fits a GAM with an AR(1) correlation structure.
- `fit_mlm`: Fits a multi-level linear model (random intercept per subject).
- `sim_study`: Ties everything together—generates data, applies dropout, then fits the chosen model and returns a tidy summary.

```{r}
#| label: sim-functions

#' Generate Synthetic Data
#'
#' This function simulates synthetic panel data for `n` participants over `n_days` time points,
#' with an intervention effect, random intercepts and slopes, and AR(1)-correlated residuals.
#'
#' @param n Number of participants. Default is 100.
#' @param n_days Number of time points per participant. Default is 28.
#' @param b Fixed effect for the intervention slope. Default is 0.1.
#' @param phi AR(1) autocorrelation coefficient. Default is 0.7.
#' @param sigma AR(1) residual standard deviation. Default is 0.6.
sim_data <- function(n = 100,
                     n_days = 28,
                     
                     # effect parameters
                     b = 3.7, # effect in unstandardized units
                     mu = 78.5, # grand mean of the outcome
                     
                     
                     # random effects parameters
                     tau_int = 9.7,   # Random intercept SD (between-person variance)
                     tau_slope = .05,  # Random slope SD 
                     within_person_sd = 11.8, # Within-person SD
                     
                     # AR(1) parameters
                     phi = 0.7,     # Autocorrelation coefficient
                     effect_shape = "grow",
                     k = .5)     
{
  dat <- tibble(
    id = 1:n,
    age = sample(18:36, n, replace = TRUE),
    gender = sample(c("man","woman","non-binary"), n, prob = c(.45, .45, .1), replace = TRUE),
    condition = factor(sample(c("control", "intervention"), n, replace = TRUE)),
    experimental_condition = ifelse(condition == "intervention", 1, 0),
    intercept = rnorm(n, 0, tau_int),
    slope = rnorm(n, 0, tau_slope)
  ) |> 
    # expand to 28 waves per id
    crossing(
      day = 1:n_days
    ) |> 
    mutate(
      intervention_period = as.numeric(day > 7 & day < 22),
      effect_time = case_when(
        effect_shape == "plateau" ~ if_else(intervention_period == 1, (b + slope) * (1-exp(-k * (day - 7))), 0),
        effect_shape == "grow" ~ if_else(intervention_period == 1, (day - 7) * ((b + slope)/7), 0),
        TRUE ~ NA_real_
      )
    ) |> 
    group_by(id) |> 
    mutate(
      sigma = within_person_sd * sqrt(1-phi^2),
      # Generate AR(1) errors for each participant
      e = as.numeric(arima.sim(n = n_days, 
                               model = list(ar = phi), 
                               sd = sigma)),
      # Add random effect + fixed effect + AR(1) error
      wellbeing = 
        mu + 
        intercept + 
        effect_time * experimental_condition * intervention_period + 
        .01*(age-18) +
        -.05*gender %in% c("women","non-binary") +
        e
    ) |> 
    ungroup()

}


#' Simulate Dropout
#'
#' Introduces missingness and dropout into a dataset by randomly assigning records as missing
#' or dropped out. Once a participant is dropped out, all subsequent records become missing.
#'
#' @param dat A tibble generated by \code{sim_data()}.
#'
#' @return A tibble of the same structure as \code{dat}, but with some \code{wellbeing} values set to NA.
#'
sim_dropout <- function(dat) {
  
  dropout <- dat |> 
    mutate(
      missing = sample(c(TRUE, FALSE), n(), replace = TRUE, prob = c(.10, .90)),
      dropout = sample(c(TRUE, FALSE), n(), replace = TRUE, prob = c(.01, .99))
    ) |>
    mutate(
      missing = ifelse(cumsum(dropout) > 0, TRUE, missing),
      .by = id
    ) |>
    arrange(as.integer(id), day) |> 
    mutate(wellbeing = ifelse(missing, NA, wellbeing))
  dropout
}

#' Fit a Generalized Additive Model (GAM)
#'
#' Fits a GAM model to the provided dataset using \code{mgcv::gam}, including an AR(1)
#' correlation structure and random intercept for each ID.
#'
#' @param dat A tibble of repeated-measures data (e.g., from \code{sim_data()} and \code{sim_dropout()}).
#'
#' @return An object of class \code{gam}, which is the fitted GAM model.
#'
fit_gam <- function(dat) {
  
  gam(wellbeing ~ 
        condition:intervention_period + age + gender +
        s(id, bs = "re") + 
        s(day, by = condition, bs = "tp"), 
      data = dat,
      correlation = corAR1(form = ~ day | id))
}

#' Fit a Multi-Level Model (MLM)
#'
#' Fits a linear mixed-effects model (LME) with random intercept for each ID using \code{lme4::lmer}.
#'
#' @param dat A tibble of repeated-measures data (e.g., from \code{sim_data()} and \code{sim_dropout()}).
#'
#' @return An object of class \code{lmerMod}, which is the fitted MLM model.
#'
fit_mlm <- function(dat) {
  # lmer(wellbeing ~ condition*intervention_period + age + gender + (1|id), data = dat)
  lme(
    fixed = wellbeing ~ condition*intervention_period + age + gender,
    random = ~ 1|id,  # or use a more flexible structure if needed
    correlation = corAR1(form = ~ day | id),
    method = "ML",
    data = dat |> filter(!is.na(wellbeing))
  )
}

fit_mlm2 <- function(dat) {
  tmp <- dat |> 
    group_by(id) |> 
    # take the mean of days 1-7 
    mutate(baseline = mean(wellbeing[day < 8], na.rm = TRUE)) |> 
    filter(intervention_period == 1) |> 
    filter(!is.na(wellbeing))
  
  lme(
    fixed = wellbeing ~ baseline + condition + age + gender,
    random = ~ 1 + condition|id,
    correlation = corAR1(form = ~ day | id),
    method = "ML",
    data = tmp |> filter(!is.na(wellbeing))
  )
}

#' Simulation Study Orchestrator
#'
#' A higher-level function that ties together data simulation, dropout, and model fitting,
#' returning a tidy summary of the fitted model parameters.
#'
#' @param model_function A function to fit the model. Defaults to \code{fit_gam}.
#' @param n Number of participants passed to \code{sim_data()}. Default is 1000.
#' @param n_days Number of time points per participant passed to \code{sim_data()}. Default is 28.
#' @param b Fixed effect for the intervention slope passed to \code{sim_data()}. Default is 0.01.
#' @param phi AR(1) autocorrelation coefficient passed to \code{sim_data()}. Default is 0.7.
#' @param sigma AR(1) residual standard deviation passed to \code{sim_data()}. Default is 0.6.
#'
#' @return A data frame (tibble) of model estimates from \code{broom::tidy(parametric = TRUE)}.
#'
sim_study <- function(model_function = fit_gam, ...) {
  args <- list(...)
  dat <- do.call(sim_data, args) |> 
    sim_dropout()
  model_function(dat) |> 
    broom::tidy(parametric = TRUE)
}

```


### Test Data Creation & Descriptively Plot
Below, we create a sample dataset using sim_data() and examine it with basic plots. The first is a smoothed trend over days for each condition; the second is a line plot of average wellbeing per day, grouped by condition.

```{r}
#| label: test-plot-descriptive

dat <- sim_data(effect_shape = "plateau")

(sds <- dat |> 
  group_by(id) |> 
  summarise(mean_value = mean(wellbeing, na.rm = TRUE),
            sd_within  = sd(wellbeing, na.rm = TRUE)) |>
  summarise(between_sd   = sd(mean_value, na.rm = TRUE),
            avg_within_sd = mean(sd_within, na.rm = TRUE)))

# line plot
dat |> 
  group_by(condition, day) |> 
  summarise(wellbeing = mean(wellbeing)) |> 
  ggplot(aes(y = wellbeing, x = day, color = condition)) +
  geom_line() + 
  theme_minimal() +
  scale_y_continuous(limits = c(60, 100))

# loess
ggplot(dat, aes(y = wellbeing, x = day, color = condition)) +
  geom_smooth() + 
  theme_minimal()


```

### Test Fitting a Model

We fit our GAM (`fit_gam`) to the newly simulated data, then demonstrate a simple way to visualize model predictions using `plot_predictions` from the marginaleffects package.

```{r}
#| label: test-fit

dat <- sim_data() |> 
  sim_dropout()

dat2 <- dat |> 
  group_by(id) |> 
  # take the mean of days 1-7 
  mutate(baseline = mean(wellbeing[day < 8], na.rm = TRUE)) |> 
  filter(intervention_period == 1)

mlm <- fit_mlm(dat)
summary(mlm)

mlm2 <- fit_mlm2(dat)
summary(mlm2)

gam <- fit_gam(dat)
plot_predictions(gam, condition = c("day", "condition"))

```

# Power analysis: Repeated Simulation Studies

To assess power/sensitivity, we run multiple simulations (controlled by `n_sims`) and gather the parameter estimates for a particular term (here, "conditionintervention:intervention_periodTRUE"). Each iteration calls `sim_study`, which does the data generation, dropout, and fitting.

```{r}
#| label: sim-studies

n_sims <- 500

specs <- expand_grid(
  model = c("fit_gam", "fit_mlm", "fit_mlm2"),  # model names as strings
  b = c(1.2, 2.4, 3.6, 4.8, 6),
  effect_shape = c("grow", "plateau")
) |> 
  (\(d) { d$row_id <- pmap_chr(d, ~ paste0(names(list(...)), "=", c(...), collapse = "_")); d })() |> 
  mutate(i = row_number())

results <- specs |> 
  rowwise() |> 
  mutate(sim_results = list({
    message("Processing spec row: ", i)
    map_dfr(1:n_sims, function(sim) {
      tryCatch({
        sim_study(
          model_function = get(model),  # retrieve the function from its name
          n = 100,
          n_days = 28,
          # effect parameters
          b = b, # effect in unstandardized units
          mu = 78.5, # grand mean of the outcome
          effect_shape = effect_shape,
          k = .5,
          # random effects parameters
          tau_int = 9.7,   # Random intercept SD (between-person variance)
          tau_slope = .8,  # Random slope SD 
          within_person_sd = 11.8, # Within-person SD
          # AR(1) parameters
          phi = 0.7     # Autocorrelation coefficient
        ) |> 
          filter(term == "conditionintervention:intervention_period" | 
                   (model == "fit_mlm2" & term == "conditionintervention")) |> 
          mutate(
            conf.low = estimate - 1.96 * std.error,
            conf.high = estimate + 1.96 * std.error,
            sim = sim
          )
      }, error = function(e) {
        message("Simulation ", sim, " failed: ", e$message)
        tibble(
          term = NA_character_,
          estimate = NA_real_,
          std.error = NA_real_,
        )
      })
    }, .progress = TRUE)
  })) |> 
  ungroup() |> 
  unnest(sim_results)

sim_summary <- results |> 
  group_by(row_id) %>%
  summarise(
    model = first(model),
    b = first(b),
    effect_shape = first(effect_shape),
    mean_effect = mean(estimate, na.rm = TRUE),
    mean_se = mean(std.error, na.rm = TRUE),
    mean_conf.low = mean(conf.low, na.rm = TRUE),
    mean_conf.high = mean(conf.high, na.rm = TRUE),
    power = sum(conf.low > 0, na.rm = TRUE) / sum(!is.na(conf.low))
  )


```

```{r}
#| label: visualize-power

# Estimated effect vs. true effect (b)
ggplot(sim_summary, aes(x = b, y = mean_effect, color = model)) +
  geom_point(size = 3) +
  geom_errorbar(aes(ymin = mean_conf.low, ymax = mean_conf.high), width = 0.1) +
  facet_wrap(~ effect_shape) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray") +
  labs(x = "True Effect (unstandardized b)", y = "Estimated Effect",
       title = "Estimated vs. True Effects by Model and Effect Shape") +
  scale_x_continuous(breaks = c(1.2, 2.4, 3.6, 4.8, 6),
                     sec.axis = sec_axis(~ . / 12, name = "Standardized Effect (b/12)"))

# Power vs. true effect (b)
ggplot(sim_summary, aes(x = b, y = power, color = model)) +
  geom_line(size = 1) +
  geom_point(size = 3) +
  facet_wrap(~ effect_shape) +
  labs(x = "True Effect (unstandardized b)", y = "Power",
       title = "Power by Model and Effect Shape") +
  scale_x_continuous(breaks = c(1.2, 2.4, 3.6, 4.8, 6),
                     sec.axis = sec_axis(~ . / 12, name = "Standardized Effect (b/12)"))
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, 1), breaks = seq(0, 1, .1))

# results |>
#   forestplot(mean = estimate,
#              lower = conf.low,
#              upper = conf.high,
#              labeltext = term)

```


# Alternative GAM specifications

```{r}
#| label: alternative-gams

# fit2 <- gam(wellbeing ~ condition*intervention_period + s(day, id, bs = "fs", m = 1), method = "REML", data = dat)
# 
# fit3 <- gamm4(wellbeing ~ condition*intervention_period + id + s(day, by = condition),
#              family = gaussian(),
#              random = ~(1|id),
#              data = dat)
# 
# ggpredict(fit, c("day","condition")) %>% plot()

```

